---
title: "Final Project"
author: " Grace Malonga, Jingyi Liang, Xu Chen"
date: "2019/12/6"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##### Preparetion

### library package
```{r lib}
library(tidyverse)
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager",repos = '"https://cloud.r-project.org')
library("BiocManager")
library(limma)
library(table1)
library(genefilter)
library(ggrepel)

```

### loaddata
```{r}
ILD <- read.csv("ILD_proteomics_data.csv")
#mutate age_stratify variable used for question 4
ILD <- ILD %>% mutate(age_stratify=ifelse(Age>=60,"old","young"))
```

##### Question 1 What proteins are differentially expressed between ILD phenotypes?
rowttest
```{r}
ild2 <- ILD %>%
 filter(Group=="RA_ILD"|Group=="RA_noILD")%>%
 mutate(Group=recode(factor(Group),RA_noILD="ra.no.ild", RA_ILD="ra.ild"))


proteins1 <- ild2 %>%
  mutate_at(vars(20:1341), funs(log2(.))) %>%
  select(20:1341) %>%
  t()

m <- as.matrix(proteins1)
f <- factor(c(rep("ra.no.ild", 37), rep("ra.ild", 39)))
ttests <- rowttests(m, f)
pval <- ttests$p.value
sum(pval<0.01)
sum(pval<0.01 & abs(ttests$dm) >= .5)
sum(pval<0.01 & ttests$dm >= .5)
ttests

names(ttests)

#mask <- with(ttests, abs(dm) >= .5 & p.value < .01) #when applying a stricter threshold with
mask <- with(ttests,p.value < .01)
spike <- rownames(proteins1) %in% colnames(proteins1)
cols <- ifelse(mask, "red", ifelse(spike,"dodgerblue","black"))

with(ttests, plot(-dm, -log10(p.value), cex = .8, pch = 16,
                  xlim=c(-2,2), ylim=c(0,5),
                  xlab="Log2 Fold Change",
                  col=cols))
abline(h=1.5, v=c(-1,1), lty=2)
title('A. T-tests Results Volcano Plot: RA ILD vs. RA no ILD')
```
```{r}
ild3 <- ILD %>% filter(Group=="IPF"|Group=="healthy")%>%
  mutate(Group=recode(factor(Group),IPF="IPF",healthy="healthy"))

proteins1 <- ild3 %>%
  mutate_at(vars(20:1341), funs(log2(.))) %>%
  select(20:1341) %>%
  t()

m <- as.matrix(proteins1)
f <- factor(c(rep("IPF", 42), rep("healthy", 42)))
ttests <- rowttests(m, f)
pval <- ttests$p.value
sum(pval<0.01)
sum(pval<0.01 & abs(ttests$dm) >= .5)
sum(pval<0.01 & ttests$dm >= .5)
ttests

names(ttests)

#mask <- with(ttests, abs(dm) >= .5 & p.value < .01) #when applying a stricter threshold with
mask <- with(ttests,p.value < .01)
spike <- rownames(proteins1) %in% colnames(proteins1)
cols <- ifelse(mask, "red", ifelse(spike,"dodgerblue","black"))

with(ttests, plot(-dm, -log10(p.value), cex = .8, pch = 16,
                  xlim=c(-2,2), ylim=c(0,5),
                  xlab="Log2 Fold Change",
                  col=cols))
abline(h=1.5, v=c(-1,1), lty=2)
title('B. T-tests Results Volcano Plot: IPF vs. Healthy')
```

```{r}
ild4 <- ILD %>% filter(Group=="RA_ILD"|Group=="IPF")%>%
  mutate(Group=recode(factor(Group),RA_ILD="RA_ILD",IPF="IPF"))

proteins1 <- ild4 %>%
  mutate_at(vars(20:1341), funs(log2(.))) %>%
  select(20:1341) %>%
  t()

m <- as.matrix(proteins1)
f <- factor(c(rep("IPF", 42), rep("RA_ILD", 39)))
ttests <- rowttests(m, f)
pval <- ttests$p.value
sum(pval<0.01)
sum(pval<0.01 & abs(ttests$dm) >= .5)
sum(pval<0.01 & ttests$dm >= .5)
ttests

names(ttests)

#mask <- with(ttests, abs(dm) >= .5 & p.value < .01) #when applying a stricter threshold with
mask <- with(ttests,p.value < .01)
spike <- rownames(proteins1) %in% colnames(proteins1)
cols <- ifelse(mask, "red", ifelse(spike,"dodgerblue","black"))

with(ttests, plot(-dm, -log10(p.value), cex = .8, pch = 16,
                  xlim=c(-2,2), ylim=c(0,5),
                  xlab="Log2 Fold Change",
                  col=cols))
abline(h=1.5, v=c(-1,1), lty=2)
title('C. T-tests Results Volcano Plot: RA ILD vs. IPF')
```

```{r}
ild5 <- ILD %>% filter(ILD_subtype=="Fibrotic NSIP"|ILD_subtype=="Cellular NSIP")%>%
  mutate(Group=recode(factor(ILD_subtype),RA_ILD="Fibrotic NSIP",IPF="Cellular NSIP"))

proteins1 <- ild5 %>%
  mutate_at(vars(20:1341), funs(log2(.))) %>%
  select(20:1341) %>%
  t()

m <- as.matrix(proteins1)
f <- factor(c(rep("Fibrotic NSIP", 19), rep("Cellular NSIP", 15)))
ttests <- rowttests(m, f)
pval <- ttests$p.value
sum(pval<0.01)
sum(pval<0.01 & abs(ttests$dm) >= .5)
sum(pval<0.01 & ttests$dm >= .5)
ttests

names(ttests)

#mask <- with(ttests, abs(dm) >= .5 & p.value < .01) #when applying a stricter threshold with
mask <- with(ttests,p.value < .01)
spike <- rownames(proteins1) %in% colnames(proteins1)
cols <- ifelse(mask, "red", ifelse(spike,"dodgerblue","black"))

with(ttests, plot(-dm, -log10(p.value), cex = .8, pch = 16,
                  xlim=c(-2,2), ylim=c(0,5),
                  xlab="Log2 Fold Change",
                  col=cols))
abline(h=1.5, v=c(-1,1), lty=2)
title('D. T-tests Results Volcano Plot: Fibrotic NSIP vs. Cellular NSIP')
```



##### Question 2 How many of these differentially expressed proteins would you expect to be false positives (think of your significance level)? Can you reduce the false positive rate? (hint: consider corrections for multiple hypothesis testing)  



### I.RA_noILD and RA_ILD
```{r}
group_list <- factor(c(rep("RA_noILD",37),rep("RA_ILD",39)))
ILD %>% filter(Group=='RA_noILD'|Group=='RA_ILD') -> data1
data1 <- data1 %>% mutate(Group=recode(factor(Group),RA_noILD="RA_noILD",RA_ILD="RA_ILD"))
data <- data1 %>% select(20:1341) %>% log2(.) %>% t()
design <- model.matrix(~0+group_list)

colnames(design) <- levels(factor(group_list))
rownames(design) <- colnames(data)

contrast.matrix<-makeContrasts(RA_noILD-RA_ILD,levels = design)
contrast.matrix
fit <- lmFit(data,design)

fit2 <- contrasts.fit(fit, contrast.matrix) 

fit2 <- eBayes(fit2,trend=TRUE)

result_limma1 <- topTable(fit2,coef=1,adjust.method='fdr',number=nrow(data))

dif1 <- result_limma1%>%filter(adj.P.Val<0.01& abs(logFC)>0.5)
dif1
#The number of differentially expressed proteins
sum(result_limma1$adj.P.Val < 0.01&abs(result_limma1$logFC)>=0.5)
sum(result_limma1$adj.P.Val < 0.01&result_limma1$logFC<=-0.5)
sum(result_limma1$adj.P.Val < 0.01&result_limma1$logFC>=0.5)
```

```{r}
result_limma1$threshold <-
  factor(ifelse(result_limma1$adj.P.Val < 0.01 & abs(result_limma1$logFC) >= 0.5,
              ifelse(result_limma1$logFC>= 0.5,
                      'Up','Down'),'NoSignifi'),levels=c('Up','Down','NoSignifi'))                         
ggplot(result_limma1,aes(x=logFC,y=-log10(adj.P.Val),color=threshold))+
  geom_point()+
   scale_color_manual(values=c("#DC143C","#00008B","#808080"))+
  theme_bw()+
  theme(
    legend.title = element_blank())+
  ylab('-log10 (p-adj)')+
  xlab('log2 (FoldChange)')+ggtitle("                  The volcano plot of RA_noILD versus RA_ILD")+
  geom_vline(xintercept=c(-0.5,0.5),lty=3,col="black",lwd=0.5) +
  geom_hline(yintercept = -log10(0.01),lty=3,col="black",lwd=0.5)
```


### II.healthy and IPF
```{r}
group_list <- factor(c(rep("healthy",42),rep("IPF",42)))
ILD %>% filter(Group=='IPF'|Group=='healthy') -> data2
data2 <- data2 %>% mutate(Group=recode(factor(Group),healthy="healthy",IPF="IPF"))
data <- data2 %>% select(20:1341) %>% log2(.) %>% t()
design <- model.matrix(~0+factor(group_list))

colnames(design)=levels(factor(group_list))

rownames(design) <- colnames(data)

contrast.matrix<-makeContrasts(healthy - IPF,levels = design)
contrast.matrix
fit <- lmFit(data,design)

fit2 <- contrasts.fit(fit, contrast.matrix) 

fit2 <- eBayes(fit2,trend=TRUE)

result_limma2 <- topTable(fit2,coef=1,adjust.method='fdr',number=nrow(data))


dif2 <- result_limma2%>%filter(adj.P.Val<0.01& abs(logFC)>0.5)
dif2
#The number of differentially expressed proteins
sum(result_limma2$adj.P.Val < 0.01&abs(result_limma2$logFC)>=0.5)
sum(result_limma2$adj.P.Val < 0.01&result_limma2$logFC<=-0.5)
sum(result_limma2$adj.P.Val < 0.01&result_limma2$logFC>=0.5)
```

```{r}

result_limma2$threshold <-
  factor(ifelse(result_limma2$adj.P.Val < 0.01 & abs(result_limma2$logFC) >= 0.5,
                ifelse(result_limma2$logFC>= 0.5,
                       'Up','Down'),'NoSignifi'),levels=c('Up','Down','NoSignifi'))                         
ggplot(result_limma2,aes(x=logFC,y=-log10(adj.P.Val),color=threshold))+
  geom_point()+
   scale_color_manual(values=c("#00008B","#DC143C","#808080"))+
  theme_bw()+
  theme(
    legend.title = element_blank())+
  ylab('-log10 (p-adj)')+
  xlab('log2 (FoldChange)')+ggtitle("                  The volcano plot of Healthy versus IPF")+
  geom_vline(xintercept=c(-0.5,0.5),lty=3,col="black",lwd=0.5) +
  geom_hline(yintercept = -log10(0.01),lty=3,col="black",lwd=0.5)
```


### III.RA_ILD and IPF 
```{r}
group_list <- factor(c(rep("RA_ILD",39),rep("IPF",42)))
ILD %>% filter(Group=='IPF'|Group=='RA_ILD') -> data3
data3 <- data3 %>% mutate(Group=recode(factor(Group),RA_ILD="RA_ILD",IPF="IPF"))
data <- data3 %>% select(20:1341) %>% log2(.) %>% t()
design <- model.matrix(~0+factor(group_list))

colnames(design)=levels(factor(group_list))

rownames(design) <- colnames(data)

contrast.matrix<-makeContrasts(RA_ILD - IPF,levels = design)
contrast.matrix
fit <- lmFit(data,design)

fit2 <- contrasts.fit(fit, contrast.matrix) 

fit2 <- eBayes(fit2,trend=TRUE)

result_limma3 <- topTable(fit2,coef=1,adjust.method='fdr',number=nrow(data))

dif3 <- result_limma3%>%filter(adj.P.Val<0.01& abs(logFC)>0.5)
dif3
#The number of differentially expressed proteins
sum(result_limma3$adj.P.Val < 0.01&abs(result_limma3$logFC)>=0.5)
sum(result_limma3$adj.P.Val < 0.01&result_limma3$logFC<=-0.5)
sum(result_limma3$adj.P.Val < 0.01&result_limma3$logFC>=0.5)
```

```{r}
result_limma3$threshold <-
  factor(ifelse(result_limma3$adj.P.Val < 0.01 & abs(result_limma3$logFC) >= 0.5,
                ifelse(result_limma3$logFC>= 0.5,
                       'Up','Down'),'NoSignifi'),levels=c('Up','Down','NoSignifi'))                         
ggplot(result_limma3,aes(x=logFC,y=-log10(adj.P.Val),color=threshold))+
  geom_point()+
   scale_color_manual(values=c("#808080","#DC143C","#00008B"))+
  theme_bw()+
  theme(
    legend.title = element_blank())+
  ylab('-log10 (p-adj)')+
  xlab('log2 (FoldChange)')+ggtitle("                  The volcano plot of RA_ILD versus IPF")+
  geom_vline(xintercept=c(-0.5,0.5),lty=3,col="black",lwd=0.5) +
  geom_hline(yintercept = -log10(0.01),lty=3,col="black",lwd=0.5)


```


### IV.RA cellular and RA fibrotic
```{r}
group_list <- factor(c(rep("CellularNSIP",15),rep("FibroticNSIP",19)))
ILD %>% filter(ILD_subtype=='Fibrotic NSIP'|ILD_subtype=='Cellular NSIP') -> data4
data4 <- data4 %>% mutate(ILD_subtype=recode(factor(ILD_subtype),FibroticNSIP="Fibrotic NSIP",CellularNSIP="Cellular NSIP")) 
data <- order(data4[,"ILD_subtype"])
data <- data4[data,]
data <- data %>% select(20:1341) %>% log2(.) %>% t()
design <- model.matrix(~0+group_list)
colnames(design) <- levels(factor(group_list))
rownames(design) <- colnames(data)

contrast.matrix<-makeContrasts(CellularNSIP - FibroticNSIP,levels = design)
contrast.matrix
fit <- lmFit(data,design)

fit2 <- contrasts.fit(fit, contrast.matrix) 

fit2 <- eBayes(fit2,trend=TRUE)

result_limma4 <- topTable(fit2,coef=1,adjust.method='fdr',number=nrow(data))

dif4 <- result_limma4%>%filter(adj.P.Val<0.01& abs(logFC)>0.5)
dif4

#The number of differentially expressed proteins
sum(result_limma4$adj.P.Val < 0.01&abs(result_limma4$logFC)>=0.5)
sum(result_limma4$adj.P.Val < 0.01&result_limma4$logFC<=-0.5)
sum(result_limma4$adj.P.Val < 0.01&result_limma4$logFC>=0.5)
```

```{r}
result_limma4$threshold <-
  factor(ifelse(result_limma4$adj.P.Val < 0.01 & abs(result_limma4$logFC) >= 0.5,
                ifelse(result_limma4$logFC>= 0.5,
                       'Up','Down'),'NoSignifi'),levels=c('Up','Down','NoSignifi'))                         
ggplot(result_limma4,aes(x=logFC,y=-log10(adj.P.Val),color=threshold))+
  geom_point()+
   scale_color_manual(values=c("#808080","#00008B","#DC143C"))+
  theme_bw()+
  theme(
    legend.title = element_blank())+
  ylab('-log10 (p-adj)')+
  xlab('log2 (FoldChange)')+ggtitle("                  The volcano plot of RA cellular versus RA fibrotic")+ 
  geom_vline(xintercept=c(-0.5,0.5),lty=3,col="black",lwd=0.5) +
  geom_hline(yintercept = -log10(0.01),lty=3,col="black",lwd=0.5)
```















#####Question 3 Are there differences in demographic, body size, or lung function variables between groups?




###Gender
```{r}
data <- ILD %>% select(Group,Gender)
dataTable <- table(data)
mosaicplot(dataTable,shade=TRUE,color=TRUE,main='Mosaic Plot of Gender',xlab='Gender',ylab='Group')
chisq.results <- chisq.test(dataTable,correct = FALSE)
chisq.results$expected

```









###Age
```{r}
boxplot(ILD$Age~ILD$Group,main='Distribution of Age',ylab='Age',xlab='Group Type')
shapiro.test(ILD$Age[ILD$Group=="healthy"])
shapiro.test(ILD$FVC[ILD$Group=="RA_noILD"])
```



###Race
```{r}
data <- ILD %>% select(Group,Race)
dataTable <- table(data)
mosaicplot(dataTable,shade=TRUE,color=TRUE,main='Mosaic Plot of Race',xlab='Race',ylab='Group')
chisq.results <- chisq.test(dataTable, simulate.p.value = TRUE)
chisq.results$expected

```





###BMI
```{r}
boxplot(ILD$BMI~ILD$Group,main='Distribution of BMI',ylab='BMI',xlab='Group Type')
```




###EverSmoker
```{r}
data <- ILD %>% select(Group,EverSmoker)
dataTable <- table(data)
mosaicplot(dataTable,shade=TRUE,color=TRUE,main='Mosaic Plot of EverSmoker',xlab='EverSmoker',ylab='Group')
chisq.results <- chisq.test(dataTable,correct = FALSE)
chisq.results$expected

```






###PkYr
```{r}
boxplot(ILD$PkYr~ILD$Group,main='Distribution of Pack years of smoking',ylab='PkYr',xlab='Group Type')
```

###FEV1
```{r}
boxplot(ILD$FEV1~ILD$Group,main='Distribution of Forced Expiratory Volume',ylab='Forced Expiratory Volume',xlab='Group Type')
```



###FEV1PP
```{r}
boxplot(ILD$FEV1PP~ILD$Group,main='Distribution of Percent of Forced Expiratory Volume Predicted',ylab='FEV1PP',xlab='Group Type')
```


###FVC
```{r}

boxplot(ILD$FVC~ILD$Group,main='Distribution of Forced Vital Capacity',ylab='Forced Vital Capacity',xlab='Group Type')
```


###FVCPP
```{r}
boxplot(ILD$FVCPP~ILD$Group,main='Distribution of Percent of Forced Vital Capacity Predicted',ylab='FVCPP',xlab='Group Type')
```



###RatioPP
```{r}
boxplot(ILD$RatioPP~ILD$Group,main='Distribution of Percent of FEV1/FVC ratio',ylab='FVCPP',xlab='Group Type')
```


###TLC
```{r}
boxplot(ILD$TLC~ILD$Group,main='Distribution of Total Lung Capacity',ylab='Total Lung Capacity',xlab='Group Type')
```


###TLCPP
```{r}
boxplot(ILD$TLCPP~ILD$Group,main='Distribution of Percent of Total Lung Capacity',ylab='TLCPP',xlab='Group Type')
```






**Emphysema**
```{r}
boxplot(ILD$Emphysema~ILD$Group,main='Distribution of Goddard score for amount of Emphysema',ylab='Emphysema',xlab='Group Type')
```


###X6MWD
```{r}
boxplot(ILD$X6MWD~ILD$Group,main='Distribution of 6 Minute Walk Distance',ylab='6 minute walk distance',xlab='Group Type')
```






RA_ILD & RA_noILD

```{r}
data1<- ILD%>%filter(Group=="RA_ILD"|Group=="RA_noILD")%>%
  mutate(Group=recode(factor(Group),RA_ILD="0",RA_noILD="1"),
         Group_1=ifelse(Group==1,1,0))
  
```

```{r}
data1$Group <-factor(data1$Group,levels=c(0,1,2),labels=c("RA_ILD","RA_noILD","P.Value"))
label(data1$PkYr)<-"Pack years of smoking"
label(data1$FEV1)<-"Forced expiratory volume"
label(data1$FEV1PP)<-"Percent of FEV1"
label(data1$FVC)<-"Forced vital capacity"
label(data1$FVCPP)<-"Percent of FCV "
label(data1$RatioPP)<-"Percent of FEV1/FVC ratio"
label(data1$TLC)<-"Total lung capacity"
label(data1$TLCPP)<-"Percent of TLC"
label(data1$Emphysema)<-"Goddard score for amount of emphysema"
label(data1$X6MWD)<-"6 minute walk distance"
units(data1$Age)<-"years"
units(data1$BMI)<-"kg/m^2"
getPvalue <- function(x, name, ...) {
    if (length(x) == 0) {
        y <- data1[[name]]
        s <- rep("", length(render.default(x=y, name=name, ...)))
        if (name=="Age") {
            p <- t.test(y ~ data1$Group_1)$p.value
        } 
        else if(name=="EverSmoker"|name=="Gender") {
            p <- chisq.test(table(y,droplevels(data1$Group)),correct = FALSE)$p.value}
        else if(name=="Race"){
          p <- chisq.test(table(y,droplevels(data1$Group)),simulate.p.value = TRUE)$p.value
        }
        else{
          p<- wilcox.test(data1[[name]] ~ data1$Group_1)$p.value
        }
        s[2] <- sub("<", "&lt;", format.pval(p, digits=3,eps=0.001))
        s
    } else {
        render.default(x=x, name=name, ...)
    }
}

pvaluelabel <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}

table1(~Gender+Age+Race+BMI+EverSmoker+PkYr+FEV1+FEV1PP+FVC+FVCPP+RatioPP+TLC+TLCPP+Emphysema+X6MWD|Group,data=data1,droplevels = F,render=getPvalue,render.strat=pvaluelabel,overall=F)

```



RA_ILD&IPF
```{r}
data1<- ILD%>%filter(Group=="IPF"|Group=="RA_ILD")%>%
  mutate(Group=recode(factor(Group),IPF="0",RA_ILD="1"),
         Group_1=ifelse(Group==1,1,0))
  
```


```{r}
data1$Group <-factor(data1$Group,levels=c(0,1,2),labels=c("IPF","RA_ILD","P.Value"))
label(data1$PkYr)<-"Pack years of smoking"
label(data1$FEV1)<-"Forced expiratory volume"
label(data1$FEV1PP)<-"Percent of FEV1"
label(data1$FVC)<-"Forced vital capacity"
label(data1$FVCPP)<-"Percent of FCV "
label(data1$RatioPP)<-"Percent of FEV1/FVC ratio"
label(data1$TLC)<-"Total lung capacity"
label(data1$TLCPP)<-"Percent of TLC"
label(data1$Emphysema)<-"Goddard score for amount of emphysema"
label(data1$X6MWD)<-"6 minute walk distance"
units(data1$Age)<-"years"
units(data1$BMI)<-"kg/m^2"
getPvalue <- function(x, name, ...) {
    if (length(x) == 0) {
        y <- data1[[name]]
        s <- rep("", length(render.default(x=y, name=name, ...)))
        if (name=="Age") {
            p <- t.test(y ~ data1$Group_1)$p.value
        } 
        else if(name=="EverSmoker"|name=="Gender") {
            p <- chisq.test(table(y,droplevels(data1$Group)),correct = FALSE)$p.value}
        else if(name=="Race"){
          p <- chisq.test(table(y,droplevels(data1$Group)),simulate.p.value = TRUE)$p.value
        }
        else{
          p<- wilcox.test(data1[[name]] ~ data1$Group_1)$p.value
        }
        s[2] <- sub("<", "&lt;", format.pval(p, digits=3,eps=0.001))
        s
    } else {
        render.default(x=x, name=name, ...)
    }
}

pvaluelabel <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}

table1(~Gender+Age+Race+BMI+EverSmoker+PkYr+FEV1PP+FVCPP+RatioPP+TLCPP+X6MWD|Group,data=data1,droplevels = F,render=getPvalue,render.strat=pvaluelabel,overall=F)

```





healthy&IPF
```{r}
data1<- ILD%>%filter(Group=="IPF"|Group=="healthy")%>%
  mutate(Group=recode(factor(Group),IPF="0",healthy="1"),
         Group_1=ifelse(Group==1,1,0))
  
```


```{r}
data1$Group <-factor(data1$Group,levels=c(0,1,2),labels=c("IPF","healthy","P.Value"))
label(data1$PkYr)<-"Pack years of smoking"
label(data1$FEV1)<-"Forced expiratory volume"
label(data1$FEV1PP)<-"Percent of FEV1"
label(data1$FVC)<-"Forced vital capacity"
label(data1$FVCPP)<-"Percent of FCV "
label(data1$RatioPP)<-"Percent of FEV1/FVC ratio"
label(data1$TLC)<-"Total lung capacity"
label(data1$TLCPP)<-"Percent of TLC"
label(data1$Emphysema)<-"Goddard score for amount of emphysema"
label(data1$X6MWD)<-"6 minute walk distance"
units(data1$Age)<-"years"
units(data1$BMI)<-"kg/m^2"
getPvalue <- function(x, name, ...) {
    if (length(x) == 0) {
        y <- data1[[name]]
        s <- rep("", length(render.default(x=y, name=name, ...)))
        if (name=="Age") {
            p <- t.test(y ~ data1$Group_1)$p.value
        } 
        else if(name=="EverSmoker"|name=="Gender") {
            p <- chisq.test(table(y,droplevels(data1$Group)),correct = FALSE)$p.value}
        else if(name=="Race"){
          p <- chisq.test(table(y,droplevels(data1$Group)),simulate.p.value = TRUE)$p.value
        }
        else{
          p<- wilcox.test(data1[[name]] ~ data1$Group_1)$p.value
        }
        s[2] <- sub("<", "&lt;", format.pval(p, digits=3,eps=0.001))
        s
    } else {
        render.default(x=x, name=name, ...)
    }
}

pvaluelabel <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}

table1(~Gender+Age+Race+EverSmoker|Group,data=data1,droplevels = F,render=getPvalue,render.strat=pvaluelabel,overall=F)

```








CellularNSIP & FibroticNSIP

```{r}
data1<- ILD%>%filter(ILD_subtype=="Cellular NSIP"|ILD_subtype=="Fibrotic NSIP")%>%
  mutate(ILD_subtype=recode(factor(ILD_subtype),"Cellular NSIP"="0","Fibrotic NSIP"="1"),
         ILD_subtype_1=ifelse(ILD_subtype==1,1,0))
  
```

```{r}
data1$ILD_subtype <-factor(data1$ILD_subtype,levels=c(0,1,2),labels=c("CellularNSIP","FibroticNSIP","P.Value"))
label(data1$PkYr)<-"Pack years of smoking"
label(data1$FEV1)<-"Forced expiratory volume"
label(data1$FEV1PP)<-"Percent of FEV1"
label(data1$FVC)<-"Forced vital capacity"
label(data1$FVCPP)<-"Percent of FCV "
label(data1$RatioPP)<-"Percent of FEV1/FVC ratio"
label(data1$TLC)<-"Total lung capacity"
label(data1$TLCPP)<-"Percent of TLC"
label(data1$Emphysema)<-"Goddard score for amount of emphysema"
label(data1$X6MWD)<-"6 minute walk distance"
units(data1$Age)<-"years"
units(data1$BMI)<-"kg/m^2"
getPvalue <- function(x, name, ...) {
    if (length(x) == 0) {
        y <- data1[[name]]
        s <- rep("", length(render.default(x=y, name=name, ...)))
        if (name=="Age") {
            p <- wilcox.test(data1[[name]] ~ data1$ILD_subtype_1)$p.value
        } 
        else if(name=="EverSmoker"|name=="Gender") {
            p <- chisq.test(table(y,droplevels(data1$ILD_subtype)),correct = FALSE)$p.value}
        else if(name=="Race"){
          p <- chisq.test(table(y,droplevels(data1$ILD_subtype)),simulate.p.value = TRUE)$p.value
        }
        else{
          p<- wilcox.test(data1[[name]] ~ data1$ILD_subtype_1)$p.value
        }
        s[2] <- sub("<", "&lt;", format.pval(p, digits=3,eps=0.001))
        s
    } else {
        render.default(x=x, name=name, ...)
    }
}

pvaluelabel <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}

table1(~Gender+Age+Race+BMI+EverSmoker+PkYr+FEV1+FEV1PP+FVC+FVCPP+RatioPP+TLC+TLCPP+Emphysema+X6MWD|ILD_subtype,data=data1,droplevels = F,render=getPvalue,render.strat=pvaluelabel,overall=F)

```

##### Question 4 Does adjustment for relevant demographic, body size, or lung function variables change your protein signature profiles? 




### I.RA_noILD vs RA_ILD young

```{r}
group_list <- factor(c(rep("RA_noILD",24),rep("RA_ILD",11)))
ILD %>% filter(Group=='RA_noILD'|Group=='RA_ILD') -> data5
data5 <- data5 %>% mutate(Group=recode(factor(Group),RA_noILD="RA_noILD",RA_ILD="RA_ILD"))
data <- data5 %>% filter(age_stratify=="young") %>% select(20:1341) %>% log2(.) %>% t()
design <- model.matrix(~0+group_list)
colnames(design) <- levels(factor(group_list))
rownames(design) <- colnames(data)
contrast.matrix<-makeContrasts(RA_noILD-RA_ILD,levels = design)
contrast.matrix                             
fit <- lmFit(data,design)

fit2 <- contrasts.fit(fit, contrast.matrix) 

fit2 <- eBayes(fit2,trend=TRUE)

result_limma5 <- topTable(fit2,coef=1,adjust.method='fdr',number=nrow(data))
dif5 <- result_limma5%>%filter(adj.P.Val<0.01& abs(logFC)>0.5)
dif5
#The number of differentially expressed proteins
sum(result_limma5$adj.P.Val < 0.01&abs(result_limma5$logFC)>=0.5)
sum(result_limma5$adj.P.Val < 0.01&result_limma5$logFC<=-0.5)
sum(result_limma5$adj.P.Val < 0.01&result_limma5$logFC>=0.5)
```

```{r}
result_limma5$threshold <-
  factor(ifelse(result_limma5$adj.P.Val < 0.01 & abs(result_limma5$logFC) >= 0.5,
                ifelse(result_limma5$logFC>= 0.5,
                       'Up','Down'),'NoSignifi'),levels=c('Up','Down','NoSignifi'))                         
ggplot(result_limma5,aes(x=logFC,y=-log10(adj.P.Val),color=threshold))+
  geom_point()+
   scale_color_manual(values=c("#DC143C","#00008B","#808080"))+
  theme_bw()+
  theme(
    legend.title = element_blank())+
  ylab('-log10 (p-adj)')+
  xlab('log2 (FoldChange)')+ggtitle("         The volcano plot of RA_noILD versus RA_ILD in young group")+
  geom_vline(xintercept=c(-0.5,0.5),lty=3,col="black",lwd=0.5) +
  geom_hline(yintercept = -log10(0.01),lty=3,col="black",lwd=0.5)
```



### RA_noILD vs RA_ILD old

```{r}
group_list <- factor(c(rep("RA_noILD",13),rep("RA_ILD",28)))
ILD %>% filter(Group=='RA_noILD'|Group=='RA_ILD') -> data6
data6 <- data6 %>% mutate(Group=recode(factor(Group),RA_noILD="RA_noILD",RA_ILD="RA_ILD"))
data <- data6 %>% filter(age_stratify=="old") %>% select(20:1341) %>% log2(.) %>% t()
design <- model.matrix(~0+group_list)
colnames(design) <- levels(factor(group_list))
rownames(design) <- colnames(data)
contrast.matrix<-makeContrasts(RA_noILD-RA_ILD,levels = design)
contrast.matrix
fit <- lmFit(data,design)

fit2 <- contrasts.fit(fit, contrast.matrix) 

fit2 <- eBayes(fit2,trend=TRUE)

result_limma6 <- topTable(fit2,coef=1,adjust.method='fdr',number=nrow(data))

dif6 <- result_limma6%>%filter(adj.P.Val<0.01& abs(logFC)>0.5)
dif6
#The number of differentially expressed proteins
sum(result_limma6$adj.P.Val < 0.01&abs(result_limma6$logFC)>=0.5)
sum(result_limma6$adj.P.Val < 0.01&result_limma6$logFC<=-0.5)
sum(result_limma6$adj.P.Val < 0.01&result_limma6$logFC>=0.5)
```

```{r}
result_limma6$threshold <-
  factor(ifelse(result_limma6$adj.P.Val < 0.01 & abs(result_limma6$logFC) >= 0.5,
                ifelse(result_limma6$logFC>= 0.5,
                       'Up','Down'),'NoSignifi'),levels=c('Up','Down','NoSignifi'))                         
ggplot(result_limma6,aes(x=logFC,y=-log10(adj.P.Val),color=threshold))+
  geom_point()+
   scale_color_manual(values=c("#808080","#DC143C","#00008B"))+
  theme_bw()+
  theme(
    legend.title = element_blank())+
  ylab('-log10 (p-adj)')+
  xlab('log2 (FoldChange)')+ggtitle("         The volcano plot of RA_noILD versus RA_ILD in old group")+
  geom_vline(xintercept=c(-0.5,0.5),lty=3,col="black",lwd=0.5) +
  geom_hline(yintercept = -log10(0.01),lty=3,col="black",lwd=0.5)
```







